A Mechanical Formalization of Higher-Ranked Polymorphic Type Inference

### 2019-03-02 main.tex
newpage before appendix; move measures to appendix; new ==>> symbol
Remove Abstract for submission

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
Modern functional programming languages, such as Haskell or OCaml,
use sophisticated forms of type inference. While an
important topic in the Programming Languages research, there is little
work on the mechanization of the metatheory of type inference in
theorem provers. In particular we are unaware of any complete
formalization of the type inference algorithms that are the backbone of modern 
functional languages.

This paper presents the first full mechanical formalization of the metatheory for
higher-ranked polymorphic type inference. The system that we formalize
is the bidirectional type system by Dunfield and
Krishnaswami (DK). The DK type system has two variants (a declarative
and an algorithmic one) that have been \emph{manually} proven
\emph{sound}, \emph{complete} and \emph{decidable}. We present 
a mechanical formalization in the Abella theorem prover 
of DK's declarative type system with a novel algorithmic system. We have
a few reasons to use a new algorithm. Firstly, our new algorithm
employs \emph{worklist judgments}, which allow precise capture of the
scope of variables and simplify the formalization of scoping in a
theorem prover. Secondly, while DK's original formalization comes with very
well-written manual proofs, there are several details missing and some
incorrect proofs,
which complicate the task of writing a formal proof. 
Despite the use of a different algorithm we prove the
same results as DK, although with significantly different proofs and
proof techniques. Since such type inference algorithms are quite
subtle and have a complex metatheory, mechanical
formalizations are an important advance in type-inference research.
\end{abstract}

@@

%% Keywords
%% comma separated list
\keywords{type inference, higher-rank polymorphism, mechanisation}  %% \keywords are mandatory in final camera-ready submission

@@

%% Appendix
\appendix
\input{sections/Appendix.tex}

>>>>

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
%\begin{abstract}
%Modern functional programming languages, such as Haskell or OCaml,
%use sophisticated forms of type inference. While an
%important topic in the Programming Languages research, there is little
%work on the mechanization of the metatheory of type inference in
%theorem provers. In particular we are unaware of any complete
%formalization of the type inference algorithms that are the backbone of modern 
%functional languages. 
%
%This paper presents the first full mechanical formalization of the metatheory for
%higher-ranked polymorphic type inference. The system that we formalize
%is the bidirectional type system by Dunfield and
%Krishnaswami (DK). The DK type system has two variants (a declarative
%and an algorithmic one) that have been \emph{manually} proven
%\emph{sound}, \emph{complete} and \emph{decidable}. We present 
%a mechanical formalization in the Abella theorem prover 
%of DK's declarative type system with a novel algorithmic system. We have
%a few reasons to use a new algorithm. Firstly, our new algorithm
%employs \emph{worklist judgments}, which allow precise capture of the
%scope of variables and simplify the formalization of scoping in a
%theorem prover. Secondly, while DK's original formalization comes with very
%well-written manual proofs, there are several details missing and some
%incorrect proofs,
%which complicate the task of writing a formal proof. 
%Despite the use of a different algorithm we prove the
%same results as DK, although with significantly different proofs and
%proof techniques. Since such type inference algorithms are quite
%subtle and have a complex metatheory, mechanical
%formalizations are an important advance in type-inference research.
%\end{abstract}

@@

%% Keywords
%% comma separated list
\keywords{type inference, higher-rank polymorphism, mechanization}  %% \keywords are mandatory in final camera-ready submission

@@

%% Appendix
\newpage
\appendix
\input{sections/Appendix.tex}
